---
    output:
      html_document:
              
        toc: true
        toc_float: false
        toc_depth: 3
        number_sections: true
        
        code_folding: show
        code_download: true
        
        fig_width: 9 
        fig_height: 4
        fig_align: "center"
        
        highlight: pygments
        theme: cerulean
        
        
    title: "Cats"
    subtitle: "All about cats"
    author: "SM"
---

```{r}
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    tidy = FALSE, # cleaner code printing
    size = "small", # smaller code
    
    fig.path = "figures/", #graphics location
    out.width = "100%",

    message = FALSE,
    warning = FALSE
    )

```

# cat

## happy cat {.tabset .tabset-fade .tabset-pills}

happy cats are happy.

```{r j}

x <- c(2,4,6)
y <- c(1,3,7)
plot(y~x)
lines(x,y)

```

### green cat {-} 

green cats are green

### white cat {-}

white cats are white

### blue cat {-}

blue cats are blue

### red cat {-}

red cats are red

# cat

## hungry cat {-}

hungry cats are hungry.

```{r}

x <- seq(-4, 4, length.out = 100)

y <- dnorm(x)

plot(x, y, type = "l", lwd = 2, col = "orange")

```

<details><summary>click here</summary>
More cat

```{r code}
# print
print("orange cat")
```


<script>
$(document).ready(function() {

  $chunks = $('.fold');

  $chunks.each(function () {

    // add button to source code chunks
    if ( $(this).hasClass('s') ) {
      $('pre.r', this).prepend("<div class=\"showopt\">Show Source</div><br style=\"line-height:22px;\"/>");
      $('pre.r', this).children('code').attr('class', 'folded');
    }

    // add button to output chunks
    if ( $(this).hasClass('o') ) {
      $('pre:not(.r)', this).has('code').prepend("<div class=\"showopt\">Show Output</div><br style=\"line-height:22px;\"/>");
      $('pre:not(.r)', this).children('code:not(r)').addClass('folded');

      // add button to plots
      $(this).find('img').wrap('<pre class=\"plot\"></pre>');
      $('pre.plot', this).prepend("<div class=\"showopt\">Show Plot</div><br style=\"line-height:22px;\"/>");
      $('pre.plot', this).children('img').addClass('folded');

    }
  });

  // hide all chunks when document is loaded
  $('.folded').css('display', 'none')

  // function to toggle the visibility
  $('.showopt').click(function() {
    var label = $(this).html();
    if (label.indexOf("Show") >= 0) {
      $(this).html(label.replace("Show", "Hide"));
    } else {
      $(this).html(label.replace("Hide", "Show"));
    }
    $(this).siblings('code, img').slideToggle('fast', 'swing');
  });
});
</script>


<style>
.showopt {
  background-color: #004c93;
  color: #FFFFFF; 
  width: 100px;
  height: 20px;
  text-align: center;
  vertical-align: middle !important;
  float: right;
  font-family: sans-serif;
  border-radius: 8px;
}

.showopt:hover {
    background-color: #dfe4f2;
    color: #004c93;
}

pre.plot {
  background-color: white !important;
}
</style>


<div class="fold s o">

```{r cars}
summary(cars)
```

</div>

enter image description here
</details>




